---
title: "Pitcher Comparison"
format:
  html:
    css: styles.css
---

## The Impact of First Pitch Strike Mentality on All Outcomes

We analyzed three distinct starting pitchers from the 2024 season:

1.  **The Sniper:** The pitcher with the **Highest** First Pitch Strike rate.
2.  **The Average:** A pitcher representing the league median.
3.  **The Wild Card:** The pitcher with the **Lowest** First Pitch Strike rate.

Does throwing that first strike actually change the composition of how their season goes?




::: {.callout-tip collapse="true" appearance="simple" icon=false}
## ℹ️ **Click here: Why is "The Sniper" important?**

**The Strategy:** This chart represents a pitcher who dominates the strike zone early (70% FPS).

**What to look for:**
* **Massive Green Slices:** Because he gets ahead 0-1, batters are forced to protect the plate, leading to high Strikeout and Field Out numbers.
* **Tiny Yellow Slice:** He rarely walks people because he doesn't fall behind in the count.
* **The Takeaway:** This is the model of efficiency. By attacking early, he minimizes damage and controls the pace of the game.
:::
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-height: 5
#| fig-width: 7

library(plotly)
library(tidyverse)

# 1. Load Data (renamed to pitcher_data to avoid errors)
pitcher_data <- read_csv("pitcher_comparison.csv")

# 2. Prepare Data
plot_data <- pitcher_data |> 
  filter(Category == "Highest FPS%") |> 
  mutate(
    fps_label = scales::percent(fps_pct, accuracy = 0.1),
    chart_title = paste0("The Sniper: ", player_name, " (", fps_label, " FPS)"),
    
    # Map Colors
    Color = case_when(
      Outcome_Type == "strikeouts" ~ "#00ff00", # Green
      Outcome_Type == "field_outs" ~ "#008000", # Dark Green
      Outcome_Type == "walks"      ~ "#ffff00", # Yellow
      Outcome_Type == "base_hits"  ~ "#ff0000", # Red
      Outcome_Type == "homeruns"   ~ "#ff00ff"  # Magenta
    ),
    
    # Clean Labels
    Clean_Label = case_when(
      Outcome_Type == "strikeouts" ~ "Strikeout",
      Outcome_Type == "field_outs" ~ "Field Out",
      Outcome_Type == "walks"      ~ "Walk",
      Outcome_Type == "base_hits"  ~ "Base Hit",
      Outcome_Type == "homeruns"   ~ "Home Run"
    )
  )

# 3. Create Plot
fig <- plot_ly(plot_data, 
        labels = ~Clean_Label, 
        values = ~Count, 
        type = 'pie',
        textinfo = 'percent',          
        hoverinfo = 'label+value',     
        marker = list(colors = ~Color, line = list(color = '#000000', width = 2)),
        textfont = list(family = "Arial", size = 14, color = "black")
)

# 4. Style (Green Title for Sniper)
fig <- fig |> layout(
  title = list(
    text = unique(plot_data$chart_title),
    font = list(color = "#00ff00", size = 20, family = "Arial") # Bright Green
  ),
  paper_bgcolor = "black",
  plot_bgcolor = "black",
  legend = list(font = list(color = "white", size = 12)),
  margin = list(t = 50, b = 20, l = 20, r = 20)
)

fig
```




::: {.callout-note collapse="true" appearance="simple" icon=false}
## ℹ️ **Click here: Understanding "The Average"**

**The Strategy:** This represents the league median. This pitcher lands a first-pitch strike about 60% of the time.

**What to look for:**
* **Balance:** You see a healthy mix of green (outs) and red/yellow (danger).
* **The Baseline:** This chart serves as your "Control Group." It shows what a typical MLB performance looks like.
* **The Takeaway:** Compare the other two charts to this one. You will clearly see how deviating from this baseline (either better or worse command) dramatically shifts the color balance.
:::
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-height: 5
#| fig-width: 7

library(plotly)
library(tidyverse)

# 1. Load Data
pitcher_data <- read_csv("pitcher_comparison.csv")

# 2. Prepare Data
plot_data <- pitcher_data |> 
  filter(Category == "Average FPS%") |> 
  mutate(
    fps_label = scales::percent(fps_pct, accuracy = 0.1),
    chart_title = paste0("The Average: ", player_name, " (", fps_label, " FPS)"),
    
    Color = case_when(
      Outcome_Type == "strikeouts" ~ "#00ff00",
      Outcome_Type == "field_outs" ~ "#008000",
      Outcome_Type == "walks"      ~ "#ffff00",
      Outcome_Type == "base_hits"  ~ "#ff0000",
      Outcome_Type == "homeruns"   ~ "#ff00ff"
    ),
    
    Clean_Label = case_when(
      Outcome_Type == "strikeouts" ~ "Strikeout",
      Outcome_Type == "field_outs" ~ "Field Out",
      Outcome_Type == "walks"      ~ "Walk",
      Outcome_Type == "base_hits"  ~ "Base Hit",
      Outcome_Type == "homeruns"   ~ "Home Run"
    )
  )

# 3. Create Plot
fig <- plot_ly(plot_data, 
        labels = ~Clean_Label, 
        values = ~Count, 
        type = 'pie',
        textinfo = 'percent',          
        hoverinfo = 'label+value',     
        marker = list(colors = ~Color, line = list(color = '#000000', width = 2)),
        textfont = list(family = "Arial", size = 14, color = "black")
)

# 4. Style (Cyan Title for Average)
fig <- fig |> layout(
  title = list(
    text = unique(plot_data$chart_title),
    font = list(color = "#00ffff", size = 20, family = "Arial") # Cyan
  ),
  paper_bgcolor = "black",
  plot_bgcolor = "black",
  legend = list(font = list(color = "white", size = 12)),
  margin = list(t = 50, b = 20, l = 20, r = 20)
)

fig
```


::: {.callout-important collapse="true" appearance="simple" icon=false}
## ℹ️ **Click here: The Danger of "The Wild Card"**

**The Strategy:** This pitcher has the lowest FPS rate (below 50%). He constantly falls behind 1-0.

**What to look for:**
* **The Yellow Danger Zone:** Notice how much larger the Walk slice (Yellow) is compared to The Sniper.
* **Hidden Damage:** When batters know a pitcher can't find the zone, they sit on specific pitches, leading to harder contact (Red slices).
* **The Takeaway:** This visualizes inefficiency. Falling behind extends innings, raises pitch counts, and gives the advantage to the batter.
:::
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-height: 5
#| fig-width: 7

library(plotly)
library(tidyverse)

# 1. Load Data (renamed to pitcher_data)
pitcher_data <- read_csv("pitcher_comparison.csv")

# 2. Prepare Data
plot_data <- pitcher_data |> 
  filter(Category == "Lowest FPS%") |> 
  mutate(
    fps_label = scales::percent(fps_pct, accuracy = 0.1),
    chart_title = paste0("The Wild Card: ", player_name, " (", fps_label, " FPS)"),
    
    # Map Colors
    Color = case_when(
      Outcome_Type == "strikeouts" ~ "#00ff00", # Green
      Outcome_Type == "field_outs" ~ "#008000", # Dark Green
      Outcome_Type == "walks"      ~ "#ffff00", # Yellow
      Outcome_Type == "base_hits"  ~ "#ff0000", # Red
      Outcome_Type == "homeruns"   ~ "#ff00ff"  # Magenta
    ),
    
    # Clean Labels
    Clean_Label = case_when(
      Outcome_Type == "strikeouts" ~ "Strikeout",
      Outcome_Type == "field_outs" ~ "Field Out",
      Outcome_Type == "walks"      ~ "Walk",
      Outcome_Type == "base_hits"  ~ "Base Hit",
      Outcome_Type == "homeruns"   ~ "Home Run"
    )
  )

# 3. Create Plot
fig <- plot_ly(plot_data, 
        labels = ~Clean_Label, 
        values = ~Count, 
        type = 'pie',
        textinfo = 'percent',          
        hoverinfo = 'label+value',     
        marker = list(colors = ~Color, line = list(color = '#000000', width = 2)),
        textfont = list(family = "Arial", size = 14, color = "black")
)

# 4. Style (Magenta Title for Wild Card)
fig <- fig |> layout(
  title = list(
    text = unique(plot_data$chart_title),
    font = list(color = "#ff00ff", size = 20, family = "Arial") # Magenta
  ),
  paper_bgcolor = "black",
  plot_bgcolor = "black",
  legend = list(font = list(color = "white", size = 12)),
  margin = list(t = 50, b = 20, l = 20, r = 20)
)

fig
```
